Elo World, a framework for \\ benchmarking weak chess \\ engines

### 2019-03-16 trunk/chess/papers/weak.tex
Multi-round elo, quartiles, generate heatmap png for figure. Submitted versions of papers

\usepackage{url}
% \usepackage{code}
% \usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{chessboard}
\usepackage{gensymb}

@@

The problem this paper is concerned with is that basically all chess
tournaments (whether with humans or computers or both) are between
players who know how to play chess, are interested in winning their
games, and have some reasonable level of skill. This makes it hard to
give a rating to weak players: They just lose every single game and so
tend towards a rating of $-\infty$\footnote{Some organizations don't
  even let ratings fall beneath a certain level, for example, the
  lowest possible USCF rating is 100.} Even if other comparatively
weak players existed to participate in the tournament and occasionally
lose to the player under study, it may still be difficult to
understand how this cohort performs in any absolute sense. (In
contrast we have ``the highest ever human rating was 2882,'' and
``there are 5,323 players with ratings in 2200--2299'' and ``Players
whose ratings drop below 1000 are listed on the next list as
'delisted'.''~\cite{fideratings}) It may also be the case that all
weak players always lose to all strong players, making it unclear just
how wide the performance gap between the two sets is. The goal of
this paper is to expand the dynamic range of chess ratings to span
all the way from extremely weak players to extremely strong ones, while
providing canonical reference points along the way.

@@

The fun part is the various algorithms, so let's get into that. First,
some ground rules and guidelines:
\begin{itemize}
\item No resigning (forfeiting) or claiming a draw. The player will
  only be asked to make a move when there exists one, and it must
  choose a move in finite time. (In practice, most of the players are
  extremely fast, with the slowest ones using about one second of
  CPU time per move.)
\item The player can retain state for the game, and executes moves
  sequentially (for either the white or black pieces), but cannot have
  state meaningfully span games. For example, it is not permitted to
  do man-in-the-middle attacks~\cite{blind} or learn opponent's moves
  from previous rounds, or to get better at chess. The majority of
  players are actually completely stateless, just a function of type
  ${\tt position} \rightarrow {\tt move}$.
\item black and white behave the same
\item Avoid game-tree search. Minimax, alpha--beta, etc. are the
  correct way to play chess programmatically. They are well-studied
  (i.e., boring) and effective, and so not well suited to our problem.
  A less obvious issue is that they are endlessly parameterizable, for
  example the search ply; this leaves us with a million things to
  fiddle with. In any case, several traditional chess engines are
  included for comparison.
\end{itemize}

@@

@@

\describeplayer{alphabetical} Make the alphabetically first move,
using standard PGN short notation. White and black both try to move
towards A1. \asymmetric \deterministic

@@

\describeplayer{no\_i\_insist} Like \player{generous}, but be overwhelmingly
polite by trying to {\em force} the opponent to accept the gift of material.
There are three tiers: Avoid at all costs mating the opponent (and moreso
checkmating). Stalemate is polite, but it is more canonical for two
polite players to form a draw by repetition, from continually offering up
material to one another and declining it. Next, avoid situations where
the opponent can refuse our gift; among these, prefer the move where the
opponent must capture the highest value piece. Finally, prefer moves
where the expected value of the offered material (i.e.~against random play)
is highest. (This means that if there are three moves, and one captures a
rook but the others capture nothing, the value is $5/3$.) This strategy
almost never wins, but is not among the worst players, since it often forces
a draw by exchanging all its pieces.

\describeplayer{reverse\_starting} This player thinks that the board is
upside-down, and as white, tries to put its pieces where black pieces
start. This often causes some conflict since the black pieces are already
there, but can also end peacefully (Figure~\ref{fig:reversestarting}).

@@

% TODO: Cyrillic tag or toast at the end
\describeplayer{cccp} Prioritize moves that Checkmate, Check, Capture
or Push, in that order. Push means to move pieces as deep as possible
into enemy territory (without any regard for their safety). Ties are
broken deterministically by the source/destination squares, so this
one is technically asymmetric.
% \begin{otherlanguage*}{russian}
% За здоровье!
% \end{otherlanguage*}
\deterministic \asymmetric \russian

@@

\describeplayer{sym\_180} As \player{sym\_mirror\_y}, but maximize
symmetry under 180\degree\ rotation
of the board (Figure~\ref{fig:sym180}. An emergent priority is
to ``castle'' with the king and queen to ``fix'' them.

@@

@@

\describeplayer{stockfish1m} As expected, the engine's performance
increases steadily as the difficulty setting increases (without
apparently affecting the time to make moves). I don't know what it's
doing with these settings. The true way to unleash chess engines is to
give them a lot of CPU and memory to search. Since the tournament is
run simultaneously across about 60 threads\footnote{ The computer is
  the completely egregious AMD 2990WX ``Threadripper~2,'' which has 32
  cores (for 64 hardware threads) and 250 Watts of power dissipation
  at load. The torture of this CPU was part of the impetus for the
  paper.} using dozens of gigabytes of memory, and sometimes I would
play {\it Hollow Knight} while it ran, I wanted to avoid having the
performance be dependent on the scheduling environment. So here,
Stockfish is given a ``node'' budget of 1~million, hence {\sf 1m}. It
takes about one second per move when running alone, and is easily the
strongest player (type) evaluated.
\traditional

@@

\describeplayer{chessmaster.nes\_lv2} Same as
\player{chessmaster.nes\_lv1}, but increasing the ``Level Of Play'' to
``Newcomer 2.'' Chessmaster has stronger levels still, but in these
modes it will think for several NES minutes, which takes multiple
seconds to emulate---too slow for our purposes. Still much
weaker than modern engines (Figure~\ref{fig:chessmaster}).
\traditional

@@

\subsection{Diultion methods}

Even at its weakest setting, \player{stockfish0} crushes all of the
nontraditional strategies. This is not surprising, but it creates
a problem for quantifying their difference in skill (do they win
one in a million games? One in 10^{100}?). Having intermediate
players allows some Elo points to flow between the two tiers
transitively. One nice way to construct intermediate players is
by interpolation. We can interpolate between any two players,\footnote{
  Even if they are stateful! The interface separates ``please
  give me a move for the current state'' and ``the following
  move was played; please update your state,'' allowing them
  to disagree. So we can just keep two independent states for
  the two players.} but interpolating with the most canonical
player, \player{random\_move}.

% Stockfish1m_32768 loses to reverse_starting in 3 moves!
1. e4 g5 2. Ba6 f5 3. Qh5

\describeplayer{stockfish1m\_r32768}
\describeplayer{stockfish1m\_r16384}
\describeplayer{stockfish1m\_r8192}
\describeplayer{stockfish1m\_r4096}
\describeplayer{stockfish1m\_r2048}
\describeplayer{stockfish1m\_r1024}
\describeplayer{stockfish1m\_r512}
\describeplayer{stockfish1m\_r256}
\describeplayer{stockfish1m\_r64}
\describeplayer{stockfish1m\_r128}

Ideally, such a system would be robust against adding new players,
but this is probably not the case.

Interpolation, like Scoville scale.

% miraculous win for "generous" strategy as black:
% playing against "suicide king". White forces black
% to mate him:
1. a4 b5 2. d3 Nh6 3. Kd2 Ng8 4. Kc3 Nh6
5. Kd4 c6 6. Ke5 f5 7. a5 Qb6 8. Bg5 Qe3+
9. Bxe3 Na6 10. f3 Nc5 11. Ra2 Ne4 12. f4 Nd2
13. Bf2 Nf3+ 14. exf3 Ng4+ 15. Kxf5 g5 16. Nh3 e5
17. Bh4 Ba3 18. b4 e4 19. Nf2 c5 20. c4 Ba6
21. g3 Rc8 22. Be2 Kf7 23. fxe4 d5 24. Qc1 dxe4
25. Rg1 bxc4 26. Rd1 cxd3 27. Qxc5 Rhf8 28. Nd2 Bc4
29. Kxg5 Rc7 30. Bf3 Ke8 31. Qe5+ Kf7 32. Qe6+ Kg7
33. Qb6 Rxf4 34. Nb3 Bb5 35. Qb7 d2 36. Nh3 Bc1
37. Qxa7 Be2 38. Nd4 h5 39. Ra1 Nf2 40. Raxc1 Nh1
41. Rf1 d1=Q 42. Qxc7+ Kh8 43. Kg6 Bc4 44. Bxd1 e3
45. Qf7 Bb3 46. Rf3 Nf2 47. Rc6 e2 48. Rc7 Rf5
49. Rc4 Rf4 50. Nb5 e1=Q 51. Rc2 Qc3 52. g4 Rf5
53. Ng1 Qc7 54. Rcc3 Nd3 55. Rc5 Nf2 56. Nc3 Ba2
57. Qh7+

% Fate-based algorithms.
% The deterministic version is fairly boring, since the
% highest probability states are similar to the starting
% position (this might not be true for "survivalist"
% or "fatalist"? check?)

% weighted random is almost the same as random, because
% the weights tend to be very similar when treated
% additively. They perform slightly different from
% random, as so (101.5 million games):

// Good site for some computer chess programs with ELO:
// http://www.computerchess.org.uk/ccrl/404/

@@

The author would like to thank the anonymous referees
of the Special Interest Group on
B
O
V
In
Chess journal for their careful reviews.

\nocite{elo1978rating}
\nocite{topple}
// citation for knights distance:
\nocite{miller2013counting}
\nocite{watkins2017losing}

>>>>

\usepackage{url}
% \usepackage{code}
% \usepackage{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{chessboard}
\usepackage{gensymb}
\usepackage{supertabular}

@@

The problem this paper addresses is that basically all chess
tournaments (whether with humans or computers or both) are between
players who know how to play chess, are interested in winning their
games, and have some reasonable level of skill. This makes it hard to
give a rating to weak players: They just lose every single game and so
tend towards a rating of $-\infty$.\footnote{Some organizations don't
  even let ratings fall beneath a certain level, for example, the
  lowest possible USCF rating is 100.} Even if other comparatively
weak players existed to participate in the tournament and occasionally
lose to the player under study, it may still be difficult to
understand how this cohort performs in any absolute sense. (In
contrast we have ``the highest ever human rating was 2882,'' and
``there are 5,323 players with ratings in 2200--2299'' and ``Players
whose ratings drop below 1000 are listed on the next list as
'delisted'.''~\cite{fideratings}) It may also be the case that all
weak players always lose to all strong players, making it unclear just
how wide the performance gap between the two sets is. The goal of
this paper is to expand the dynamic range of chess ratings to span
all the way from extremely weak players to extremely strong ones, while
providing canonical reference points along the way.

@@

The fun part is the various algorithms, so let's get into that. First,
some ground rules and guidelines:
\begin{itemize}
\item No resigning (forfeiting) or claiming a draw. The player will
  only be asked to make a move when there exists one, and it must
  choose a move in finite time. (In practice, most of the players are
  extremely fast, with the slowest ones using about one second of
  CPU time per move.)
\item The player can retain state for the game, and executes moves
  sequentially (for either the white or black pieces), but cannot have
  state meaningfully span games. For example, it is not permitted to
  do man-in-the-middle attacks~\cite{blind} or learn opponent's moves
  from previous rounds, or to get better at chess. The majority of
  players are actually completely stateless, just a function of type
  ${\tt position} \rightarrow {\tt move}$.
\item The player should try to ``behave the same'' when playing with
  the white or black pieces. Of course this can't be literally true,
  and in fact some algorithms can't be symmetric, so it's, like, a
  suggestion.
\item Avoid game-tree search. Minimax, alpha--beta, etc. are the
  correct way to play chess programmatically. They are well-studied
  (i.e., boring) and effective, and so not well suited to our problem.
  A less obvious issue is that they are endlessly parameterizable, for
  example the search ply; this leaves us with a million things to
  fiddle with. In any case, several traditional chess engines are
  included for comparison.
\end{itemize}

@@

\begin{figure}[t]
\begin{tabular}{cl}
  \deterministic & The player is deterministic \\
  \traditional & Traditional approach to chess (e.g. engine) \\
  \vegetarian & Vegetarian or vegetarian option available \\
  \canonical & A canonical algorithm! \\
  \stateful & Stateful (not including pseudorandom pool) \\
  \asymmetric & Asymmetric \\
\end{tabular}
\caption{Key} \label{fig:key}
\end{figure}

@@

\describeplayer{alphabetical} Make the alphabetically first move,
using standard PGN short notation (e.g. ``a3'' $<$ ``O-O'' $<$ ``Qxg7''). White and black both try to move
towards A1. \asymmetric \deterministic

@@

% Looks like this is the only one with *no* wins!
\describeplayer{no\_i\_insist} Like \player{generous}, but be overwhelmingly
polite by trying to {\em force} the opponent to accept the gift of material.
There are three tiers: Avoid at all costs mating the opponent (and moreso
checkmating). Stalemate is polite, but it is more canonical for two
polite players to form a draw by repetition, from continually offering up
material to one another and declining it. Next, avoid situations where
the opponent can refuse our gift; among these, prefer the move where the
opponent must capture the highest value piece. Finally, prefer moves
where the expected value of the offered material (i.e.~against random play)
is highest. (This means that if there are three moves, and one captures a
rook but the others capture nothing, the value is $5/3$.) This strategy
almost never wins, but is not among the worst players, since it often forces
a draw by exchanging all its pieces.

\describeplayer{reverse\_starting} This player thinks that the board is
upside-down, and as white, tries to put its pieces where black pieces
start. Since here we have a specific configuration in mind, we can produce
a distance metric by computing the total distance from each piece to
its desired destination. Each piece uses a different distance metric;
Chebyshev distance for the King, Manhattan distance for the Rook, and
a pretty weird function for the Knight~\cite{miller2013counting}.
Trying to move the pieces into the reversed starting position often
causes some conflict since the black pieces are already there, but can
also end peacefully (Figure~\ref{fig:reversestarting}).

@@

\describeplayer{cccp} Prioritize moves that Checkmate, Check, Capture
or Push, in that order. Push means to move pieces as deep as possible
into enemy territory (without any regard for their safety). Ties are
broken deterministically by the source/destination squares, so this
one is technically asymmetric.
\deterministic \asymmetric \russian

@@

\describeplayer{sym\_180} As \player{sym\_mirror\_y}, but maximize
symmetry under 180\degree\ rotation
of the board (Figure~\ref{fig:sym180}). An emergent priority is
to ``castle'' with the king and queen to ``fix'' them.

@@

% (could work this into a footnote)
% weighted random is almost the same as random, because
% the weights tend to be very similar when treated
% additively. They perform slightly different from
% random, as so (101.5 million games):

@@

\describeplayer{stockfish1m} As expected, the engine's performance
increases steadily as the difficulty setting increases (without
apparently affecting the time to make moves). I don't know what it's
doing with these settings. The true way to unleash chess engines is to
give them a lot of CPU and memory to search. Since the tournament is
run simultaneously across about 60 threads\footnote{ The computer is
  the completely egregious AMD 2990WX ``Threadripper~2,'' which has 32
  cores (for 64 hardware threads) and 250 Watts of power dissipation
  at load. The torture of this CPU was part of the impetus for the
  paper.} using dozens of gigabytes of memory, and sometimes I would
play {\it Hollow Knight} ({\it aka} \knight) while it ran, I wanted to avoid having the
chess skill be dependent on the scheduling environment. So here,
Stockfish is given a ``node'' budget of 1~million, hence {\sf 1m}. It
takes about one second per move when running alone, and is easily the
strongest player (type) evaluated.
\traditional

@@

\describeplayer{chessmaster.nes\_lv2} As above, but increasing the
``Level Of Play'' to ``Newcomer 2.'' Chessmaster has stronger levels
still, but in these modes it will think for several NES minutes, which
takes multiple seconds to emulate---too slow for our purposes. Still
much weaker than modern engines (Figure~\ref{fig:chessmaster}).
\traditional

@@

\subsection{Interpolation methods}

Even at its weakest setting, \player{stockfish0} crushes all of the
nontraditional strategies. This is not surprising, but it creates a
problem for quantifying their difference in skill (do they win one in
a million games? One in $10^{100}$?). Having intermediate players
allows some Elo points to flow between the two tiers transitively. One
nice way to construct intermediate players is by interpolation. We can
interpolate between any two players,\footnote{ Even if they are
  stateful! The interface separates ``please give me a move for the
  current state'' and ``the following move was played; please update
  your state,'' allowing them to disagree. So we can just keep two
  independent states for the two players.} but it is most natural to
interpolate with the most canonical player, \player{random\_move}.

\describeplayer{stockfish1m\_r{\it nnn}} There are 15 players in this
group, each characterized by a number $nnn$. Before each move, we
generate a random 16-bit number; if the number is less than $nnn$ then
we play a random move and otherwise, a move as \player{stockfish1m}.
The values of $nnn$ are chosen so that we mix a power-of-two
fraction of noise: \player{stockfish1m\_r32768} blends half random
with half strong moves, but we also have $1/4$ random, $1/8$ random,
$1/16$, $1/32$, \ldots, $1/1024$. These give us a nice smooth
gradation at high levels of play (Figure~\ref{fig:heatmap}). At the
extremes, many games have no random moves, and the performance is
difficult to distinguish from \player{stockfish1m}.
% in fact, highest IS r64?
Even at half random moves, \player{stockfish1m\_r32768} consistently
beats the non-traditional players. So we also dilute stockfish by
mixing with a majority of random moves: \player{stockfish1m\_r49152}
is $3/4$ random, and we also have $7/8$, $15/16$, $31/32$, and
$63/64$. At this point it becomes hard to distinguish from
\player{random\_move}.

Note that playing $1/64$ excellent moves and the rest randomly doesn't
do much to help one's performance. On the other hand, playing one
random move for $63/64$ strong ones does create a significant
handicap! Against strong opponents, it's easy to see how a mistake can
end the game. Even against weak ones, a blunder can be fatal
(Figure~\ref{fig:s32768}).
% TODO: graph elo with random fraction

\begin{figure}[ht]
\chessboard[setfen=rnbqkbnr/ppppp2p/B7/5ppQ/4P3/8/PPPP1PPP/RNB1K1NR b KQkq - 1 3,showmover=false]
\caption{A demonstration of how quickly random play can ruin a game.
  \player{stockfish1m\_r32768} (black; plays half random moves
  and half strong moves) loses to \player{reverse\_starting} in one of
  the shortest possible sequences: 1.~e4 g5 2.~Ba6 f5 3.~Qh5++. Both
  of black's moves are bad (so they must be random) but white's 2.~Ba6
  is a blunder as well. White is just pushing pieces as far as possible;
  the mate is purely accidental.} \label{fig:s32768}
\end{figure}

%  \describeplayer{stockfish1m\_r64512} 63/64
%  \describeplayer{stockfish1m\_r63488} 31/32
%  \describeplayer{stockfish1m\_r61440} 15/16
%  \describeplayer{stockfish1m\_r57344} 7/8
%  \describeplayer{stockfish1m\_r49152} 3/4
%  \describeplayer{stockfish1m\_r32768} 1/2
%  \describeplayer{stockfish1m\_r16384} 1/4
%  \describeplayer{stockfish1m\_r8192} 1/8
%  \describeplayer{stockfish1m\_r4096} 1/16
%  \describeplayer{stockfish1m\_r2048} 1/32
%  \describeplayer{stockfish1m\_r1024} 1/64
%  \describeplayer{stockfish1m\_r512} 1/128
%  \describeplayer{stockfish1m\_r256} 1/256
%  \describeplayer{stockfish1m\_r128} 1/512
%  \describeplayer{stockfish1m\_r64} 1/1024

\section{Running the tournament}

Given the players, it's a simple matter to simulate a bunch of games.
Since the tournament runs for tens of thousands of CPU hours, there is
of course some code to allow it to checkpoint its work, to make
efficient use of CPU and RAM, and to provide attractive ANSI graphics
showing off its activity---but this is all straightforward.\footnote{
  The source code can be found at
  \url{sourceforge.net/p/tom7misc/svn/HEAD/tree/trunk/chess/}
} The
tournament accumulates win/loss/draw counts for each pair of players
(playing as both white and black), as well as example games used to
debug or create the figures in this paper.

After the tournament is done, or from any checkpoint, I can run a
separate program to produce Elo (and other) rankings. Elo depends
on the order in which games are played, because it is designed to take
into account changes in player skill over time. However, it is easy
to just make up a random order in which the games were played, because
these players do not change over time.

Despite its strong reputation, I found Elo to be rather finnicky. It
is sensitive to the starting scores for the players, and the factor
$k$ which governs how strong of a correction to make after a game.
With players of such vastly different skill, it is also very sensitive
to the order in which the games are played.\footnote{For example,
  suppose the weak player \player{random\_move} has one win against
  \player{stockfish1m}. If this win happens early, when the two
  players have their initial Elo scores, then it doesn't affect
  anything. If it happens last, when \player{stockfish1m} has
  thousands of Elo points over \player{random\_move}, then it produces
  a very strong correction.} For similar reasons, it is also very
sensitive to imbalance in the number of games played between a
particular pair of players; if some player mostly has games against
weak players, this can artificially inflate its score.

To control for these effects, I compute the maximum number $n$ for
which we have $n$ games played between every pair of players. I then
seed each player with an initial Elo score of 1000, and sample exactly
$n$ games to play from each cell without replacement. This is so that
we play a balanced number of games. I also randomize the global order
of all games. Then I do this over again, with a smaller update factor
$k$, for 20 passes. I don't know whether it's the fact that I reduce
$k$ over time, or that I end up with effectively 20 times the number
of games played,\footnote{To be clear, running 20 passes means that
  games can be reused, which is not ideal.} but doing this
significantly reduces variance. I run the whole thing 19 times and the
result is the median Elo score. I also computed the 25\th\ and
75\th\ percentiles, which were within 1--2\% of the median, which is
good. The Elo scores appear in Section~\ref{sec:results}; $n > 400$
for this run.

Ideally, such a system would be robust against adding new players, but
this is probably not the case; it is easy to see from the results
(Figure~\ref{fig:heatmap}) that there are certain matchups that favor
one of the players despite its average skill. During development, I
often noticed significant swings in Elo scores as players were added,
especially before there were middle-tier players in the mix. One way
to deal with this would be to run the Elo simulations multiple times
while randomly ablating players from the tournament; we could then at
least estimate the variance in the Elo scores under the removal of
players, which is like adding players in reverse. I did not implement
such a thing because of oppressive SIGBOVIK deadlines.

\section{Results} \label{sec:results}

% maybe whole-page?
\begin{figure}[ht]
  \includegraphics[width=\linewidth]{elo-heatmap}
  \caption{ The matrix of outcomes for all players. There is too much
    data to print, so we just have a bunch of colors, mostly because
    it looks rad. If you don't have a color printer or TV it will
    probably not look rad. Rows indicate the player playing as white,
    from worst (\player{worstfish}, top) to best
    (\player{stockfish1m}, bottom). Columns for the player with the
    black pieces, in the same order from left to right. Green
    indicates a cell that is predominately wins (for white), red is
    losses, and blue is draws. The right and bottom third of the
    graphic are all different levels/dilutions of strong engines
    (Stockfish and Topple). This creates a nice smooth gradient where
    better engines regularly beat even slightly weaker ones, but start
    to produce draws at the highest levels. They consistently destroy
    weak engines; a cell with an $\times$ indicates that it only
    contained wins or only contained losses (not even draws).
    \\
    At the low end, there is a large splat of matchups that mostly
    produce draws against one another, but can consistently beat
    the weakest tier. In the top-left corner, a square of blue
    draws from low-ambition play; these aggressively bad players
    almost never win games, even when playing each other.
    \\
    Microtexture outside of these broad trends comes from matchups
    where the player is unusually suited or weak against that
    particular opponent. For example, the bright red cell on the
    diagonal near the center is \player{cccp} vs. \player{cccp}; this
    determinisic strategy alwasy wins in self-play as black.
  } \label{fig:heatmap}
\end{figure}

The main use of the Elo~World tournament is to benchmark some new
engine or algorithm for playing chess, particularly if it is not that
good~\cite{blind}. There are a few ways that we can interpret the
results:

The {\bf Elo score}, as described.

We can find a {\bf comparable interpolated player}. This is like the
  Scoville scale for measuring how spicy something is: Some
  golden-tongued blind tasters are given the pure chili oil and asked
  to distinguish it from a cup of sugar water, which they of course
  can do. They then dilute the chili oil 1:1 with sugar water, and try
  again. The number of dilutions before the testers cannot reliably
  tell the drink from sugar water yields the score on the Scoville
  scale. Here, for example, we can say that \player{blind\_spycheck}
  performs between \player{stockfish1m\_r63488} and
  \player{\_r61440}, so it is approximately a
  93.75--96.875\% diluted Stockfish.

We can compute a {\bf Markov probability}. The tournament table can be
  easily thought of as a Markov transition matrix.
  Imagine that there is a
  Champion trophy, held by the player corresponding to some row. Each
  cell in that row contains the probability that in a game betweeen
  those two players, the trophy would be passed to that player. We
  treat draws as choosing one of the two sides by a coin flip (or,
  equivalently, splitting the trophy in half); and for the games that
  the row player wins, the trophy is retained (probability mass assigned
  to a self-transition). It is easy to compute this matrix from the
  win/loss/draw counts in the tournament table, and it is not sensitive
  to imbalance like the Elo calculation is.
  Presented this way, we can compute the stationary distribution (if
  it exists, which it typically will), which basically tells us that
  after an extremely large number of games, what is the probability
  that a given player holds the Champion trophy? This tends to
  agree with the Elo score, but there are a few places where it does
  not (e.g. \player{same\_color} has a much higher p(Champion) score
  than its Elo seems to warrant).

And so finally, a table with numbers:
\smallskip

% \begin{footnotesize}
\tablefirsthead{%
  \hline
  Player & Elo & p(Champion) \\
  \hline
  }
\tablehead{%
  \hline
  \ldots\ Player & Elo & p(Champion) \\
  \hline
}
\tabletail{%
  \multicolumn{3}{|c|}{\ldots} \\
  \hline
}
\tablelasttail{%
  \hline
}
\begin{supertabular}{|r|l|l|}
\player{worstfish} & 188.54 & 0.00000071 \\
\player{pacifist} & 286.90 & 0.00000509 \\
\player{alphabetical} & 320.91 & 0.00000276 \\
\player{generous} & 342.71 & 0.00000306 \\
\player{popular} & 349.48 & 0.00000374 \\
\player{dangerous} & 349.60 & 0.00000291 \\
\player{safe} & 350.64 & 0.00000454 \\
\player{first\_move} & 357.71 & 0.00000434 \\
\player{rare} & 361.39 & 0.00000323 \\
\player{no\_i\_insist} & 378.23 & 0.00000244 \\
\player{huddle} & 399.55 & 0.00000933 \\
\player{sym\_180} & 429.94 & 0.00000331 \\
\player{same\_color} & 430.90 & 0.00002590 \\
\player{opposite\_color} & 432.11 & 0.00000293 \\
\player{sym\_mirror\_x} & 438.87 & 0.00000305 \\
\player{survivalist} & 439.03 & 0.00000681 \\
\player{random\_move} & 439.21 & 0.00000462 \\
\player{fatalist} & 439.89 & 0.00000314 \\
\player{sym\_mirror\_y} & 442.50 & 0.00000716 \\
\player{reverse\_starting} & 445.56 & 0.00000301 \\
\player{suicide\_king} & 447.02 & 0.00000493 \\
\player{stockfish1m\_r64512} & 462.82 & 0.00000297 \\
\player{stockfish1m\_r63488} & 482.44 & 0.00000356 \\
\player{blind\_yolo} & 488.97 & 0.00000488 \\
\player{blind\_kings} & 501.98 & 0.00000633 \\
\player{equalizer} & 504.21 & 0.00000653 \\
\player{stockfish1m\_r61440} & 521.45 & 0.00001173 \\
\player{swarm} & 534.03 & 0.00000623 \\
\player{blind\_spycheck} & 546.91 & 0.00002554 \\
\player{cccp} & 553.54 & 0.00000544 \\
\player{min\_oppt\_moves} & 597.04 & 0.00001076 \\
\player{stockfish1m\_r57344} & 600.54 & 0.00001112 \\
\player{stockfish1m\_r49152} & 752.22 & 0.00000737 \\
\player{chessmaster.nes\_lv1} & 776.15 & 0.00002055 \\
\player{stockfish1m\_r32768} & 976.20 & 0.00003025 \\
\player{chessmaster.nes\_lv2} & 989.21 & 0.00012094 \\
\player{stockfish1m\_r16384} & 1277.73 & 0.00012509 \\
\player{stockfish0} & 1335.69 & 0.00008494 \\
\player{stockfish5} & 1644.63 & 0.00070207 \\
\player{stockfish1m\_r8192} & 1690.15 & 0.00152052 \\
\player{topple10k} & 1771.65 & 0.00179298 \\
\player{stockfish10} & 1915.23 & 0.00257648 \\
\player{stockfish15} & 1952.93 & 0.00310436 \\
\player{stockfish1m\_r4096} & 2020.25 & 0.00886928 \\
\player{stockfish20} & 2139.47 & 0.00721173 \\
\player{topple1m} & 2218.71 & 0.01091286 \\
\player{stockfish1m\_r2048} & 2261.50 & 0.03132272 \\
\player{stockfish1m\_r1024} & 2425.72 & 0.06759788 \\
\player{stockfish1m\_r512} & 2521.78 & 0.11122236 \\
\player{stockfish1m\_r256} & 2581.97 & 0.15364889 \\
\player{stockfish1m\_r128} & 2609.45 & 0.17861429 \\
\player{stockfish1m\_r64} & 2637.78 & 0.20508090 \\
\player{stockfish1m} & 2644.10 & 0.21523142 \\
\end{supertabular}
% \end{footnotesize}

%  % miraculous win for "generous" strategy as black:
%  % playing against "suicide king". White forces black
%  % to mate him:
%  1. a4 b5 2. d3 Nh6 3. Kd2 Ng8 4. Kc3 Nh6
%  5. Kd4 c6 6. Ke5 f5 7. a5 Qb6 8. Bg5 Qe3+
%  9. Bxe3 Na6 10. f3 Nc5 11. Ra2 Ne4 12. f4 Nd2
%  13. Bf2 Nf3+ 14. exf3 Ng4+ 15. Kxf5 g5 16. Nh3 e5
%  17. Bh4 Ba3 18. b4 e4 19. Nf2 c5 20. c4 Ba6
%  21. g3 Rc8 22. Be2 Kf7 23. fxe4 d5 24. Qc1 dxe4
%  25. Rg1 bxc4 26. Rd1 cxd3 27. Qxc5 Rhf8 28. Nd2 Bc4
%  29. Kxg5 Rc7 30. Bf3 Ke8 31. Qe5+ Kf7 32. Qe6+ Kg7
%  33. Qb6 Rxf4 34. Nb3 Bb5 35. Qb7 d2 36. Nh3 Bc1
%  37. Qxa7 Be2 38. Nd4 h5 39. Ra1 Nf2 40. Raxc1 Nh1
%  41. Rf1 d1=Q 42. Qxc7+ Kh8 43. Kg6 Bc4 44. Bxd1 e3
%  45. Qf7 Bb3 46. Rf3 Nf2 47. Rc6 e2 48. Rc7 Rf5
%  49. Rc4 Rf4 50. Nb5 e1=Q 51. Rc2 Qc3 52. g4 Rf5
%  53. Ng1 Qc7 54. Rcc3 Nd3 55. Rc5 Nf2 56. Nc3 Ba2
%  57. Qh7+
%

% Good site for some computer chess programs with ELO:
% http://www.computerchess.org.uk/ccrl/404/

@@

The author would also like to thank the anonymous referees
of the Special Interest Group on
Bafflingly
Overdone
Ventures
In
Chess journal for their careful reviews.
